### 2025-01-13 sections/2-methods.tex
moved repeated rnn alg
cleaned up some algorithm descriptions

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $\mathcal{R}$, the root cluster
        \Require $q$, a query
        \Require $k$, the number of neighbors to find
        \State $H \Leftarrow [\ ]$, a max-heap of size $k$ to contain neighbors prioritized by distance to the query
        \State $r \Leftarrow radius$ of $\mathcal{R}$
        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \While{$\sum_{C \in Q} |C| < k$}
            \If{$Q = \emptyset$}
                \State $r \Leftarrow 2 \cdot r$
            \ElsIf{$\sum_{C \in Q} |C| >= k$}
                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
            \EndIf
            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \EndWhile
        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
        \State \textbf{return} $H$ as a list
    \end{algorithmic}
\end{algorithm}

@@

\begin{minipage}{0.5125\textwidth}
    \begin{algorithm}[H]\small
        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
        \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
        \begin{algorithmic} % enter the algorithmic environment
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find
            \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
            \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
            \Loop
                \If{$k = \sum_{(\_, \_, m) \in Q} m$}
                    \State \textbf{break}
                \EndIf
                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
                \For{$(p, \delta^{+}_{p}, m) \in Q$}
                    \If{$\delta^{-}_{p} > \tau$}
                        \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
                    \EndIf
                \EndFor
                \For{$(p, \delta^{+}_{p}, m) \in Q$}
                    \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
                    \If{$p$ is a point}
                        \State \textbf{continue}
                    \ElsIf{$p$ is a leaf}
                        \For{$c \in p$}
                            \State $Q \Leftarrow Q \cup \{ (c, \delta_{c}, 1) \}$
                        \EndFor
                    \Else
                        \State $[L, R] \Leftarrow$ children of $p$
                        \State $l, r \Leftarrow$ centers of $L, R$
                        \State $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
                        \State $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                    \EndIf
                \EndFor
            \EndLoop
            \State QuickSelect($Q$, $k$)
            \State \textbf{return} The first $k$ points in $Q$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.425\textwidth}
    \begin{algorithm}[H]\small
        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
        \label{alg:methods:depth-first-sieve}
        \begin{algorithmic}
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find

            \State{$Q \Leftarrow [\mathcal{R}]$, a min-priority queue by $\delta^{-}$}
            \State{$H \Leftarrow []$, a max-priority queue by $\delta$}

            \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
                \While{$\neg (Q.peek$ is a leaf)}
                    \State{$C \Leftarrow Q.pop$, the closest cluster}
                    \State{$[L, R] \Leftarrow$ children of $C$}
                    \State{$Q.push(L)$}
                    \State{$Q.push(R)$}
                \EndWhile
                \State{$leaf \Leftarrow Q.pop$}
                \For{$p \in leaf$}
                    \State{$H.push(p)$}
                \EndFor
                \While{$|H| > k$}
                    \State{$H.pop$}
                \EndWhile
            \EndWhile
            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

We then remove from $Q$ any element for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
Finally, we replace all remaining clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.

We continue this process until $Q$ no longer contains any clusters.
We then use the QuickSelect algorithm one last time to reorder $Q$ and return the $k$ nearest neighbors.

@@

Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
$Q$ starts containing only the root cluster $\mathcal{R}$ while $H$ starts empty.
So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:

\begin{itemize}
    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
\end{itemize}

This process terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
This leaves $H$ containing exactly the $k$ nearest neighbors to the query.

>>>>

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $\mathcal{R}$, the root cluster
        \Require $q$, a query
        \Require $k$, the number of neighbors to find
        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
        \State $r \Leftarrow radius$ of $\mathcal{R}$
        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \While{$\sum_{C \in Q} |C| < k$}
            \If{$Q = \emptyset$}
                \State $r \Leftarrow 2 \cdot r$
            \ElsIf{$\sum_{C \in Q} |C| >= k$}
                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
            \EndIf
            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
        \EndWhile
        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
        \State \textbf{return} $H$ as a list
    \end{algorithmic}
\end{algorithm}

@@

@@

\begin{minipage}{0.5125\textwidth}
    \begin{algorithm}[H]\small
        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
        \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
        \begin{algorithmic} % enter the algorithmic environment
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find
            \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
            \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
            \While{$\sum_{(\_, \_, m) \in Q} m \neq k$}
                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
                \State $Q^{'} \Leftarrow \emptyset$
                \For{$(C, \_, \_) \in Q$}
                    \If{$\delta^{-}_{C} \leq \tau$}
                        \If{$C$ is a point}
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
                        \ElsIf{$C$ is a leaf}
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
                        \Else
                            \State $[L, R] \Leftarrow$ children of $C$
                            \State $l, r \Leftarrow$ centers of $L, R$
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                        \EndIf
                    \EndIf
                \EndFor
                \State $Q \Leftarrow Q^{'}$
            \EndWhile
            \State QuickSelect($Q$, $k$)
            \State \textbf{return} The first $k$ points in $Q$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.425\textwidth}
    \begin{algorithm}[H]\small
        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
        \label{alg:methods:depth-first-sieve}
        \begin{algorithmic}
            \Require $\mathcal{R}$, the root cluster
            \Require $q$, a query
            \Require $k$, the number of neighbors to find
            \State{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
            \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
            \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
                \While{$Q.peek$ is not a leaf}
                    \State{$C \Leftarrow Q.pop$, the closest cluster}
                    \State{$[L, R] \Leftarrow$ children of $C$}
                    \State{$Q.push(L)$}
                    \State{$Q.push(R)$}
                \EndWhile
                \State{$leaf \Leftarrow Q.pop$}
                \For{$p \in leaf$}
                    \State{$H.push(p)$}
                \EndFor
            \EndWhile
            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

Next, we go over all items in $Q$, skipping over any for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
If the item corresponds to a point, we keep it.
If the item corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
If the item corresponds to a non-leaf cluster, we to $Q$ the pairs of 3-tuples corresponding to its child clusters.
We continue this process until the sum of multiplicities in $Q$ is exactly $k$.
We then use the QuickSelect algorithm one last time to reorder $Q$ and return the $k$ nearest neighbors.

@@

Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$ and the ability to maintain its maximum size) of points prioritized by $\delta$.
$Q$ starts containing only the root cluster $\mathcal{R}$ while $H$ starts empty.
So long as $H$ is not full or the top element in $H$ has $\delta$ greater than or equal to the top element in $Q$, we take the following steps:

\begin{itemize}
    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
\end{itemize}

This process terminates when $H$ is full and the top element in $H$ has $\delta$ less than the top element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
